---
title: "Demonstration report on `r {params$continent}`"
format:
  html:
    embed-resources: true
params:
  continent: "All"
execute: 
  echo: false
  message: false
  warning: false
---
```{r}
#| include: false

library("dplyr")
library("ggplot2")


# Apply afcharts theme and scale defaults globally
# Ensures consistent styling across all plots
afcharts::use_afcharts()

# setting a seed so our random choices come out the same every time.
# set to today's date so we don't have hundreds of projects using seed 1 or 42.
set.seed(20260309)

# Print digits to 2 decimals
options(digits = 2)



# Load Gapminder dataset into local object
# Used as the single data source across all tabs
gap <- gapminder::gapminder

if (params$continent != "All") {
  gap <- gap |>
    filter(continent == params$continent)
}

# filtering down to a number of countries that doesn't break the categorical palette.
countries <- gap |>
  distinct(country) |>
  slice_sample(n = 4)
```

# Life expectancy over time

To keep things relatively simple, we will limit ourselves to a sample of four countries within `r params$continent`.

```{r}
gap |>
  # using a join as a filter
  inner_join(countries, by = join_by("country")) |>
  ggplot(aes(x = year, y = lifeExp, colour = country)) +
  geom_line() +
  labs(x = NULL,
         y = "Life expectancy",
         colour = "")
  
```

# Relationship between GDP/capita and life expectancy

We now use all countries in `r params$continent`.

```{r}
if (nrow(countries) == 4) {
  gap |>
    filter(year == max(year)) |>
    ggplot(aes(x = gdpPercap, y = lifeExp)) +
    geom_point() +
    geom_smooth(method = "lm") +
    scale_x_continuous(labels = scales::label_dollar()) +
    labs(x = "GDP per capita",
           y = "Life expectancy")
} else {
  # Oceania has two countries and it breaks the line of best fit.
  gap |>
    filter(year == max(year)) |>
    ggplot(aes(x = gdpPercap, y = lifeExp)) +
    geom_point() +
    scale_x_continuous(labels = scales::label_dollar()) +
    labs(x = "GDP per capita",
           y = "Life expectancy")
}
```
```{r}
# generate the model the graph above uses, so we can do some automagical commentary
model <- lm(lifeExp ~ gdpPercap, data = gap)

# We'll use this one  a few times
slope <- model$coefficients["gdpPercap"]
pos_or_neg <- if_else(slope > 0, "positive", "negative")
increase_or_decrease <- if_else(slope > 0, "increase", "decrease")
```

We see a `r pos_or_neg` relationship between GDP per capita and life expectancy in `r params$continent`.

For each additional $1000 of GDP per capita, we see an `r increase_or_decrease` of `r abs(slope*1000)` years of life expectancy. 
